Mega-analysis of the Interoceptive Accuracy Scale (IAS) Structure and its Subjective Correlates

### 2025-11-03 paper/1_Initial/manuscript.qmd
updated manuscript

<!-- Relationship with other interoception measures -->
Regarding its validity, the IAS has naturally been compared to other interoception-related measures, and shows a positive correlations with most facets of the MAIA [@mehling2018multidimensional; @gaggero2021], except for the Not-Distracting and Not-Worrying subscales [@brand2023] - which were previously highlighted as related to non-interoceptive abilities [@ferentzi2021examining].
Interestingly, findings on the correlation between the IAS and the body awareness dimension of the BPQ, have been mixed: some studies report small positive correlations [@brand2023; @campos2021; @koike2023], while others find small negative correlations [@lin2023] or no correlation at all [@murphy2019; @gaggero2021].
Positive correlations have also been observed with the "observation" and "description" subscales of the Five Facet Mindfulness Questionnaire [FFMQ, @baer2006using; @brand2023; @koike2023], as well as with the "non-reactivity" and "acting with awareness" subscales [@koike2023].
Additionally, the IAS has shown a positive correlation with the interoceptive awareness subscale of the Eating Disorder Inventory [EDI-IA, @lin2023] and a negative correlation with the Interoceptive Confusion Questionnaire [ICQ, @brewer2016alexithymia], as reported by @brand2023 and @murphy2019.
Lastly, the correlation with the Interoceptive Attention Scale [IATS, @gabriele2022dissociations] appears rather small [@lin2023; @koike2023].

@@

<!-- Interoception and correlates - item specific -->
The IAS showed positive associations with most MAIA facets, except for the Not-Distracting and Not-Worrying subscales, replicating prior findings [@brand2023]. Associations with the BPQ Body Awareness subscale were consistent with previous reports on a non-existing association between these measures [@murphy2019; @gaggero2021], whereas indices of autonomic reactivity were more strongly and negatively associated with IAS scores, supporting earlier observations that reactivity may reflect aspects of bodily signal processing captured by self-reported interoceptive accuracy [@todd2022].

@@

Associations with personality traits followed a similar domain-specific structure. Effects for the full IAS largely reflected contributions from specific pairings: Openness showed its strongest associations for Breathing–Heart, Neuroticism and Honesty–Humility for Hungry–Thirsty and Bruise–Blood sugar, and Conscientiousness for the full scale. These results indicate that associations between interoceptive self-report and personality dimensions may depend on the bodily domain represented.

Associations with belief systems were uniformly small. Weak positive associations emerged between the IAS and certain primal world beliefs (e.g., perceiving the world as comprehensible or alive) and selected conspiracy belief dimensions, particularly Global Conspiracies and Personal Wellbeing. These effects were limited to specific pairings - most often Hungry–Thirsty and Bruise–Blood sugar—indicating that links between interoceptive self-report and broader cognitive or worldview constructs are modest and domain-specific.

>>>>

<!-- Relationship with other interoception measures -->
Regarding its validity, the IAS has naturally been compared to other interoception-related measures, and shows a positive correlations with most facets of the MAIA [@mehling2018multidimensional; @gaggero2021], except for the Not-Distracting and Not-Worrying subscales [@brand2023] - which were previously highlighted as related to non-interoceptive abilities [@ferentzi2021examining].
Interestingly, findings on the correlation between the IAS and the body awareness dimension of the BPQ, have been mixed: some studies report small positive correlations [@brand2023; @campos2021; @koike2023], while others find small negative correlations [@lin2023] or no correlation at all [@murphy2019].
Positive correlations have also been observed with the "observation" and "description" subscales of the Five Facet Mindfulness Questionnaire [FFMQ, @baer2006using; @brand2023; @koike2023], as well as with the "non-reactivity" and "acting with awareness" subscales [@koike2023].
Additionally, the IAS has shown a positive correlation with the interoceptive awareness subscale of the Eating Disorder Inventory [EDI-IA, @lin2023] and a negative correlation with the Interoceptive Confusion Questionnaire [ICQ, @brewer2016alexithymia], as reported by @brand2023 and @murphy2019.
Lastly, the correlation with the Interoceptive Attention Scale [IATS, @gabriele2022dissociations] appears rather small [@lin2023; @koike2023].

@@

<!-- Interoception and correlates - item specific -->
The IAS showed positive associations with most MAIA facets, except for the Not-Distracting and Not-Worrying subscales, replicating prior findings [@brand2023]. Associations with the BPQ Body Awareness subscale were consistent with previous reports on positive association between these measures [@brand2023; @campos2021; @koike2023; @gaggero2021], whereas indices of autonomic reactivity were more strongly and negatively associated with IAS scores, supporting earlier observations that reactivity may reflect aspects of bodily signal processing captured by self-reported interoceptive accuracy [@todd2022].

@@

Associations with personality traits followed a similar domain-specific structure. Effects for the full IAS largely reflected contributions from specific pairings: Extraversion showed its strongest associations for Breathing–Heart, Neuroticism and Honesty–Humility for Hungry–Thirsty, and Openess, Conscientiousness and Agreeableness for the full scale. These results indicate that associations between interoceptive self-report and personality dimensions may depend on the bodily domain represented.

Associations with belief systems were uniformly small. Weak positive associations emerged between the IAS and certain primal world beliefs (e.g., perceiving the world as comprehensible or alive) and selected conspiracy belief dimensions, particularly Global Conspiracies and Personal Wellbeing. These effects were limited to specific pairings, most often Hungry–Thirsty and Bruise–Blood sugar, indicating that links between interoceptive self-report and broader cognitive or worldview constructs are modest and domain-specific.
